\thetitle

### 2013-04-21 tfpie/tfpie.tex
spellcheck

Unions, and recursive unions in particular, are a fundamental kind of
data definition that students are well-versed in from the previous
semester.  A fundamental early lesson is how to represent (recursive)
unions using classes and how to write recursive methods.  As an
example, figure~\ref{fig:tree} defines binary trees of numbers
(an archetypical recursive union data definition)
using the BSL language and the Class language.
%% \begin{verbatim}
%%     ;; A list of numbers (LoN) is one of:
%%     ;; - (new empty%)
%%     ;; - (new cons% Number LoN)

@@

The structure of this data definition is analagous to the approach of
the previous semester but this example brings to light an important
difference with the functional approach.  The method for computing the
sum of the empty tree is defined in the {\tt leaf} class, while the
method for computing the sum of a node is in the {\tt node} class.
When a tree object is sent the {\tt sum} method, there is no function
with a conditional to determine whether the object is a
leaf---instead, the object itself takes care of computing the sum
based on the {\tt sum} method it contains.  This shift in perspective
is at the core of object-orientation: objects contain their own
behavior and the case analysis previously done in functions is
eliminated.

@@

Thanks to the the preparation in the first half of the course, we can
cover functional OO programming in Java in a just a few lectures.  We
then increase the subset of the language we use to encompass mutation,
loops, and mutable data structures. We present \texttt{ArrayList}s,
followed briefly by arrays. Students use, and then implement, hash
tables as well as other mutable and immutable data structures.
Conventional input and output are treated only very briefly, as we
focus instead of both fundamentals and exercises making use of real
APIs such as hashing functions or Twitter posting.  Finally, while,
for, and for-each loops are presented, following the methodology of
\htdc which connects loops to stylized use of recursive functions with
accumulators, a technique the students now have two semsesters of
practice with.

@@

However, the ``day 1'' transition from the student languages used with
\htdp to \profj is too abrupt and too large.
%
Most significantly, changing languages from the first semester
immediately rather than simply adding a new concept confuses too many
issues for students.  On the first day of a \htdc-based course,
students see object-orientation, a new programming paridigm; Java, a
new language with new syntax, and static types, a crucial but
orthogonal concept.
%
In contrast, our course presents just one of these concepts on the
first day, but  covers all of them by the end of the semester.

@@

Finally, the Java portion of the class does not continue the use of
``World''-style interactive graphical
programming. \citet{local:java-world}.  Instead, our course focuses on
coverage of standard Java libraries, as well as introductory
algorithmic and data structure topics.  Continuing to use World-style
programming in motivating examples might be valuable for continutity
between the two halves of the course.

\subsection*{Acknowledgements}

>>>>

Unions, and recursive unions in particular, are a fundamental kind of
data definition that students are well-versed in from the previous
semester.  A fundamental early lesson is how to represent (recursive)
unions using classes and how to write recursive methods.  As an
example, figure~\ref{fig:tree} defines binary trees of numbers
(an archetypal recursive union data definition)
using the BSL language and the Class language.
%% \begin{verbatim}
%%     ;; A list of numbers (LoN) is one of:
%%     ;; - (new empty%)
%%     ;; - (new cons% Number LoN)

@@

The structure of this data definition is analogous to the approach of
the previous semester but this example brings to light an important
difference with the functional approach.  The method for computing the
sum of the empty tree is defined in the {\tt leaf} class, while the
method for computing the sum of a node is in the {\tt node} class.
When a tree object is sent the {\tt sum} method, there is no function
with a conditional to determine whether the object is a
leaf---instead, the object itself takes care of computing the sum
based on the {\tt sum} method it contains.  This shift in perspective
is at the core of object-orientation: objects contain their own
behavior and the case analysis previously done in functions is
eliminated.

@@

Thanks to the the preparation in the first half of the course, we can
cover functional OO programming in Java in a just a few lectures.  We
then increase the subset of the language we use to encompass mutation,
loops, and mutable data structures. We present \texttt{ArrayList}s,
followed briefly by arrays. Students use, and then implement, hash
tables as well as other mutable and immutable data structures.
Conventional input and output are treated only very briefly, as we
focus instead of both fundamentals and exercises making use of real
APIs such as hashing functions or Twitter posting.  Finally, while,
for, and for-each loops are presented, following the methodology of
\htdc which connects loops to stylized use of recursive functions with
accumulators, a technique the students now have two semesters of
practice with.

@@

However, the ``day 1'' transition from the student languages used with
\htdp to \profj is too abrupt and too large.
%
Most significantly, changing languages from the first semester
immediately rather than simply adding a new concept confuses too many
issues for students.  On the first day of a \htdc-based course,
students see object-orientation, a new programming paradigm; Java, a
new language with new syntax, and static types, a crucial but
orthogonal concept.
%
In contrast, our course presents just one of these concepts on the
first day, but  covers all of them by the end of the semester.

@@

Finally, the Java portion of the class does not continue the use of
``World''-style interactive graphical
programming. \citet{local:java-world}.  Instead, our course focuses on
coverage of standard Java libraries, as well as introductory
algorithmic and data structure topics.  Continuing to use World-style
programming in motivating examples might be valuable for continuity
between the two halves of the course.

\subsection*{Acknowledgments}
